\subsubsection{AIX}

R has been built successfully on AIX 6.1 on `powerpc', more
specifically on a platform containing a Power6 processor, by a group
around Kurt Hornik at the Wirtschaftsuniversit\"at in Vienna. We no
longer support AIX prior to 4.2, and configure will throw an 
error on such systems. 

A small part of the software needed to build R and/or install packages
is available directly from the AIX Installation DVDs, e.g., Java6,
X11, perl. Further open source software (OSS) is packaged for AIX in so-called
.rpm files and available from
the ``AIX Toolbox for Linux Applications'' from
IBM~\url{http://www-03.ibm.com/systems/power/software/aix/linux/} as
well as on \url{http://www.oss4aix.org/download}. The latter website typically
offers more recent versions of the available OSS. All needed tools
(e.g., gcc, gfortran, libreadline, etc.)
from those repositories are typically installed to
\code{/opt/freeware} and hence the executables are found in
\code{/opt/freeware/bin}, so if using those tools ensure this is in
your path. 

The recent testing has been done under AIX 6.1 using
compilers from both the GNU Compiler Collection 4.2.4 which is
available from both of the above repositories and the IBM C (XL
C)/FORTRAN (XL F) compilers in several configurations.



32-bit versions of R could be built with configure --without-iconv as well as --enable-R-shlib. For 64-bit versions he used

     OBJECT_MODE=64
     CC="gcc -maix64"
     CXX="g++ -maix64"
     F77="gfortran -maix64"
     FC="gfortran -maix64"

and was also able to build with the IBM xlc and Hitachi f90 compilers by

     OBJECT_MODE=64
     CC="xlc -q64"
     CXX="g++ -maix64"
     F77="f90 -cpu=pwr4 -hf77 -parallel=0 -i,L -O3 -64"
     FC="f90 -cpu=pwr4 -hf77 -parallel=0 -i,L -O3 -64"
     FLIBS="-L/opt/ofort90/lib -lhf90vecmath -lhf90math -lf90"

Some systems have f95 as an IBM compiler that does not by default accept FORTRAN 77. It needs the flag -qfixed=72, or to be invoked as xlf_r.

The AIX native iconv does not support encodings `latin1' nor `""' and so cannot be used. (As far as we know GNU libiconv could be installed.) Fan Long reports success on AIX 5.3 using

     OBJECT_MODE=64
     LIBICONV=/where/libiconv/installed
     CC="xlc_r -q64"
     CFLAGS="-O -qstrict"
     CXX="xlC_r -q64"
     CXXFLAGS="-O -qstrict"
     F77="xlf_r -q64"
     AR="ar -X64"
     CPPFLAGS="-I$LIBICONV/include -I/usr/lpp/X11/include/X11"
     LDFLAGS="-L$LIBICONV/lib -L/usr/lib -L/usr/X11R6/lib"
